The Folly of Passwords

### 2014-12-13 paper.tex
More work on chapter 3.

@@

Our test desktop machine was configured to accepted login
with a smartcard reader and an OpenPGP smartcard
\parencite{pietig2009functional}. The OpenPGP card integrates
with GnuPG to provide a public-private key pair that can
be used for encryption.

GnuPG has a gpg-agent service that allows these keys to
be used for SSH authentication as well, which can provide
a Single Sign-On experience betweeen logging in to a local
desktop and connecting out to remote servers.

For desktop login, we were able to use the ``Poldi'' PAM
module. The use of the smartcard requires a PIN and
the card has a lock-out after 3 incorrect tries. This is
behaviour users can recognise from their bank smartcards
and the lock-out feature negates the guessability of
even a weak 4 to 6 digit PIN.

@@

>>>>

\cite{beautement2009compliance} introduce the idea of a
``compliance budget'' which models how users perceive these costs
and benefits. Their work suggests a cap on the effectiveness of any
security measures within an organisation, but also introduces management
approaches to raise the threshold above which a user will no longer
comply with security policies. This arguably increases the cost-benefit
ratio by making it more worthwhile for users to practice secure
behaviour (even if by increasing sanctions such that avoidance thereof becomes
more beneficial than working around costs from complying with security
requirements). They still admit there is a finite limit to improvements to
this compliance budget through management, training and cultural changes.
This supports the idea that many of the costs simply must be moved to the
organisation itself.

@@

Our test desktop machine was configured to accepted login with a smartcard
reader and an OpenPGP smartcard \parencite{pietig2009functional}. The OpenPGP
card integrates with GnuPG to provide a public-private key pair that can
be used for encryption. For desktop login, we were able to use the ``Poldi''
PAM module.

GnuPG has a gpg-agent service that allows these keys to be used for SSH
authentication as well, which can provide a Single Sign-On experience
between logging in to a local desktop and connecting out to remote servers.

The use of the
smartcard requires a PIN and the card has a lock-out after 3 incorrect tries.
This is behaviour users can recognise from their bank smartcards and the
lock-out feature negates the guessability of even a weak 4 to 6 digit PIN.

@@

The difficulty of eliminating a password or PIN from the login workflow
suggests that the PKI token or smartcard approach is more suitable for
situations where security requirements are higher and two-factor authentication
is preferable. The same technology is used for debit/credit card transactions
and for withdrawing cash from a bank account and thus we would expect it
to be used in other organisations that have similar security requirements.

The \emph{Scute} project provides a PKCS~\#11 module that allows an OpenPGP
smartcard to be used with the Network Security Services (NSS) within
the Firefox web browser (we believe Chrome also supports this, but Firefox
was used in our test network). Combined with Apache's \emph{mod_nss} (which
derives from mod_ssl), we were able to implement an authentication layer
over a web-based application that supported an OpenPGP smartcard.

NSS also allows the same key to be used within Mozilla Thunderbird for
email authentication
